---
title: "Datasets"
subtitle: "From [Lizard guts: Naxos](https://github.com/katekathrynkat/lizard-guts-naxos)"
output:
  katereR::t_analysis
---

**Kate Culhane**, Dept. of Ecology, Evolution, and Marine Biology, UC Santa Barbara  
[kathrynculhane@ucsb.edu](kathrynculhane@ucsb.edu)
[<i class="fa fa-github"></i>](https://github.com/katekathrynkat/)
[<i class="fa fa-twitter"></i>](https://twitter.com/katekathrynkat/)

***

```{r setup, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
knitr::opts_knit$set(root.dir='..')
knitr::opts_chunk$set(echo=TRUE, results='hold', warning=FALSE, message=FALSE,
                      fig.width=5, fig.height=3, rows.print=5)
```

<div class = "summary">

**SUMMARY**

Description and wrangling of datasets. Raw data are located in the project directory [`data_raw`](https://github.com/katekathrynkat/PROJECT/tree/master/data_raw/).

</div>

***

**DATA FILE STRUCTURE**

    data_raw
    ├── field                        # Field/lab measurements
        ├── original_spreadsheets        # Original data (no cleaning/proofing)
            ├── [...].xlxs                  # Original Excel files
            └── data_errata.txt             # Log of corrections made to .csv files
        ├── diet_field.csv           # Stomach pump invert ID (all islands)
        ├── naxos_meta.csv           # Site metadata (Naxos)
        ├── naxos_veg.csv            # Vegetation quadrats (Naxos)
        ├── prey_meta.csv            # Pitfall/sticky metadata (all islands)
        ├── prey_pitfall.csv         # Pitfall trap invert ID (all islands)
        └── prey_sticky.csv          # Sticky trap invert ID (all islands)

<br>

**DEPENDENCIES**

```{r dependencies, class.source='fold-show'}

##### PACKAGES #####

library(tidyverse) # data manipulation & visualization
library(lubridate) # wrangling dates
library(kableExtra) # pretty tables

##### DATA #####

# Site info
meta_raw <- read_csv("data_raw/field/naxos_meta.csv")
veg_raw <- read_csv("data_raw/field/naxos_veg.csv")

# Invertebrate traps
prey_meta_raw <- read_csv("data_raw/field/prey_meta.csv",
                          col_types = "??cccc???????????????") %>% # dates as chr
  filter(island == "Naxos")
prey_p_raw <- read_csv("data_raw/field/prey_pitfall.csv") %>% 
   filter(island == "Naxos")
prey_s_raw <- read_csv("data_raw/field/prey_sticky.csv") %>% 
   filter(island == "Naxos")

# Lizard diet
diet_raw <- read_csv("data_raw/field/diet_field.csv") %>% 
  filter(island == "Naxos")

```

<br>

**OUTPUT FILES**

- `00_diet_inverts.csv` - Invertebrate counts by order from lizard stomach pumping samples
- `00_diet_matrix.csv` - Site by species matrix (diet data)
- `00_diet_veg.csv` - Vegetation counts from lizard stomach pumping samples
- `00_inverts_all.csv` - Invertebrate counts by order from both trapping and diet samples
- `00_inverts_matrix.csv` - Site by species matrix (all invert data: trapping + diet)
- `00_prey.csv` - Invertebrate counts by order from sticky and pitfall trap samples
- `00_prey_matrix.csv` - Site by species matrix (invert trap data)
- `00_prey_metadata.csv` - Metadata for sticky and pitfall trap samples
- `00_trap_veg` - Vegetation cover by life form from quadrats associated with invert traps

## Sites

- Nine sites
- "High use" (many stone walls/terraces) or "low use" (no stone walls/terraces)
- "High veg" (high % vegetation cover) or "low veg" (low % vegetation cover)

```{r}

meta_raw %>% 
  arrange(use, veg)

```

## Vegetation surveys

- Collected from quadrats
- Variables:
  - Vegetation cover by life form (grass, forb, shrub, moss, dead, open, rock, other)
  - Average vegetation height 
  - Distance to nearest wall/terrace
  
### Trap-level vegetation

```{r}

# Trap metadata
trap_veg <- prey_meta_raw %>% 
  select(2, 8, 11:21)

write_csv(trap_veg, "output/00_trap_veg.csv", col_names = TRUE)

trap_veg

```

- One quadrat associated with each invertebrate trap pair (12 per site)
- No data for MALYKO

### Site-level vegetation

```{r}

veg_raw

```

- Randomly placed
- Eight quadrats per site

## Lizard diet

- Diet samples were collected from live lizards via stomach pumping
- Samples were identified to order and counted under a microscope
- Abundance estimates are conservative (e.g. 3 beetle elytra counted as 2 beetles)
- Total number of lizard stomach samples = `r nrow(diet_raw)`

<br>

<i class="fa fa-clipboard-list"></i> **Number of samples per site**

```{r}

diet <- diet_raw %>%
  select(where(~ any(. != 0)), -island) # remove columns with 0 observations

diet %>% 
  count(site) %>% 
  arrange(-n) %>% 
  pivot_wider(names_from = site, values_from = n) %>%
  kbl() %>%
  kable_minimal(full_width = FALSE, position = "left")

```

### Inverts

```{r}

diet_inverts <- diet %>% 
  select(-c("berry":"stick","sand")) %>% 
  mutate(egg = insect_egg + spider_egg,
         larvae = coleoptera_larvae + diptera_larvae + 
           lepidoptera_larvae + neuroptera_larvae) %>% 
  select(-c("insect_egg":"spider_egg",
            "coleoptera_larvae":"neuroptera_larvae")) %>% 
  relocate(comments, .after = last_col()) %>% 
  pivot_longer("blattodea":"larvae",
               names_to = "order", values_to = "count") %>% 
  filter(count != 0) %>% 
  mutate(sample = paste0("DI-", site, "-",
                         str_split_fixed(lizard_id, "-", 3)[,2], "_",
                         str_split_fixed(lizard_id, "-", 3)[,3]),
         sample_type = "diet") %>%
  select(sample, site, sample_type, everything()) %>% 
  relocate(comments, .after = last_col())

```

- `r length(unique(diet_inverts$lizard_id))` stomach samples containing invertebrates

<br>

<i class="fa fa-clipboard-list"></i> **Number of samples containing each order:**

```{r}

count(diet_inverts, order) %>%
  arrange(-n)%>%
  kbl() %>%
  kable_minimal(full_width = FALSE, position = "left")

```

- Removed taxa present in <5 samples from future analyses
- The pseudoscorpions really fucked with nmds plots

```{r}

diet_inverts_2 <- diet_inverts %>% 
  group_by(order) %>% 
  mutate(n = length(sample)) %>% 
  filter(n>=10) %>% 
  select(-n)

write_csv(diet_inverts_2, "output/00_diet_inverts.csv", col_names = TRUE)

```

### Site by species matrix

```{r paged.print=FALSE}

diet_matrix <- diet_inverts_2 %>% 
  select(-date, -comments) %>% 
  mutate(order = factor(order, # order factor by n
                        levels = arrange(count(diet_inverts, order),-n)$order)) %>% 
  pivot_wider(names_from = order, values_from = count) %>% 
  replace(., is.na(.), 0)

write_csv(diet_matrix, "output/00_diet_matrix.csv", col_names = TRUE)

diet_matrix

```

- For community analyses with `vegan`

### Herbivory

```{r}

diet_veg <- diet %>% 
  select(-c("blattodea":"gastropoda", "sand")) %>% 
  mutate(seed = seed + seed_pod,
         sample = paste0("DV-", site, "-",
                         str_split_fixed(lizard_id, "-", 3)[,2], "_",
                         str_split_fixed(lizard_id, "-", 3)[,3]),
         sample_type = "diet_veg") %>% 
  select(-seed_pod, -comments) %>% 
  pivot_longer("berry":"stick",
               names_to = "veg", values_to = "count") %>% 
  filter(count != 0) %>% 
  select(sample, site, sample_type, everything())

write_csv(diet_veg, "output/00_diet_veg.csv", col_names = TRUE)

diet_veg

```

- `r length(unique(diet_veg$lizard_id))` stomach samples containing vegetation

<br>

<i class="fa fa-clipboard-list"></i> **Number of samples containing each type of veg:**

```{r}

count(diet_veg, veg) %>%
  arrange(-n)%>%
  kbl() %>%
  kable_minimal(full_width = FALSE, position = "left")

```

## Invertebrate trapping

- 12 trap pairs (1 sticky trap + 1 pitfall trap) per site
  - Sticky traps = yellow cards
  - Pitfall traps = cups with antifreeze
- Traps open 33-35 hours
- Condition for each trap coded as "GOOD" (effort = 1), "FLAW" (effort = 0.5), or "LOST" (effort = 0) based on comments

### Metadata 

```{r}

# Trap metadata
prey_meta <- prey_meta_raw %>% 
  left_join(prey_p_raw[c(2,4,31)], by = c("site", "trap")) %>% 
  left_join(prey_s_raw[c(2,4,24)], by = c("site", "trap")) %>% 
  mutate(comments_p = comments.y,
         comments_s = comments,
         comments = comments.x,
         set = dmy_hm(paste(date_set, time_set),
                      quiet = TRUE), # suppress warnings
         collected = dmy_hm(paste(date_coll, time_coll),
                            quiet = TRUE), # suppress warnings
         hours_open = as.numeric(collected - set)*24,
         pitfall_effort = case_when(
           pitfall_condition == "GOOD" ~ 1,
           pitfall_condition == "FLAW" ~ 0.5,
           pitfall_condition == "LOST" ~ 0),
         sticky_effort = case_when(
           sticky_condition == "GOOD" ~ 1,
           sticky_condition == "FLAW" ~ 0.5,
           sticky_condition == "LOST" ~ 0
         )) %>%
  select(site, set, collected, hours_open, crew, trap,
         pitfall_condition, pitfall_effort, sticky_condition, sticky_effort,
         comments, comments_p, comments_s)

write_csv(prey_meta, "output/00_prey_meta.csv", col_names = TRUE)

prey_meta

```

### ID/count data

- Samples were identified to order and counted under a microscope

```{r}

# Pitfall trap data
prey_p <- prey_p_raw %>% 
  mutate(sample_type = "pitfall",
         sample = paste0("TP-", site, "-", trap)) %>% 
  select(-gecko, -lizard) %>% # remove columns for non-invertebrates
  pivot_longer(5:28, names_to = "order", values_to = "count") 

# Sticky trap data
prey_s <- prey_s_raw %>% 
  mutate(sample_type = "sticky",
         sample = paste0("TS-", site, "-", trap)) %>% 
  select(-gecko) %>% # remove columns for non-invertebrates
  pivot_longer(5:22, names_to = "order", values_to = "count")

# All trap data
prey <- full_join(prey_p, prey_s) %>% 
  filter(count != 0) %>% 
  select(sample, site, trap, sample_type, order, count)

prey

```

<br>

<i class="fa fa-clipboard-list"></i> **Number of samples containing each order:**

```{r}

count(prey, order) %>%
  arrange(-n)%>%
  kbl() %>%
  kable_minimal(full_width = FALSE, position = "left")

```

- Several orders not represented in the diet samples and so removed from further analyses

```{r}

prey_2 <- prey %>% 
  filter(order %in% unique(diet_inverts_2$order))

write_csv(prey_2, "output/00_prey.csv", col_names = TRUE)

```

### Site by species matrix

```{r paged.print=FALSE}

prey_matrix <- prey_2 %>% 
  pivot_wider(names_from = order, values_from = count) %>% 
  replace(., is.na(.), 0)

write_csv(prey_matrix, "output/00_prey_matrix.csv", col_names = TRUE)

prey_matrix

```

- For community analyses with `vegan`


## Combined dataset (trapping + diet)

### Abundance data

```{r}

inverts_all <- full_join(prey_2, diet_inverts_2) %>% 
  select(-trap, -date, -lizard_id, -comments)

write_csv(inverts_all, "output/00_inverts_all.csv", col_names = TRUE)

inverts_all

```

- Combined trapping (sticky + pitfall) and diet data

### Site by species matrix

```{r paged.print=FALSE}

inverts_matrix <- inverts_all %>% 
  mutate(order = factor(order, # order factor by n
                        levels = arrange(count(inverts_all, order),-n)$order)) %>% 
  pivot_wider(names_from = order, values_from = count) %>% 
  replace(., is.na(.), 0)

write_csv(inverts_matrix, "output/00_inverts_matrix.csv", col_names = TRUE)

inverts_matrix

```

- For community analyses with `vegan`

<br>
<br>
<br>

***

<details><summary>**SESSION INFO**</summary>
```{r session-info}
devtools::session_info()
```
</details>
