---
title: "Metadata"
subtitle: "From [Lizard guts: Naxos](https://github.com/katekathrynkat/lizard-guts-naxos)"
output:
  katereR::t_metadata
---

**Kate Culhane**, Dept. of Ecology, Evolution, and Marine Biology, UC Santa Barbara  
[kathrynculhane@ucsb.edu](kathrynculhane@ucsb.edu)
[<i class="fa fa-github"></i>](https://github.com/katekathrynkat/)
[<i class="fa fa-twitter"></i>](https://twitter.com/katekathrynkat/)

***

```{r setup, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(echo = TRUE, results = 'hold',
                      warning = FALSE, message = FALSE,
                      fig.width=5, fig.height=3, rows.print=5)
```

<div class = "summary">

**Summary**

Descriptions of raw datasets used in further analyses. Datasets are located in the project directory [`data_raw`](https://github.com/katekathrynkat/lizard-guts-naxos/data_raw/).

</div>

***

#### File locations

    data_raw
    ├── field                        # Field/lab measurements
        ├── original_spreadsheets        # 
            ├── [...].xlxs                  # 
            └── data_errata.txt             # 
        ├── .csv            

Limitations:

- Prey/diet IDed to order
- Prey measured as relative abundance

```{r packages and data, class.source = 'fold-show'}

##### PACKAGES #####

library(tidyverse) # data manipulation & visualization

##### DATA #####

# Brief description
read_csv()

```

## Field/lab data

### Dataset 1

Brief description of data

**Variables**

- `variable_name`: definition

```{r}

iris

```

## Literature data

## Spatial data

## Important things

<div class = "important">

Something important

</div>

> Note about something to check later

## Notes & resources

- [Webpage](link), author

<br>
<br>
<br>

<details><summary>**Session info**</summary>
```{r code}
devtools::session_info()
```
</details>

