---
title: "Metadata"
subtitle: "From [Lizard guts: Naxos](https://github.com/katekathrynkat/lizard-guts-naxos)"
output:
  katereR::t_metadata
---

**Kate Culhane**, Dept. of Ecology, Evolution, and Marine Biology, UC Santa Barbara  
[kathrynculhane@ucsb.edu](kathrynculhane@ucsb.edu)
[<i class="fa fa-github"></i>](https://github.com/katekathrynkat/)
[<i class="fa fa-twitter"></i>](https://twitter.com/katekathrynkat/)

***

```{r setup, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(echo = TRUE, results = 'hold',
                      warning = FALSE, message = FALSE,
                      fig.width=5, fig.height=3, rows.print=5)
```

<div class = "summary">

**Summary**

Descriptions of raw datasets used in further analyses. Datasets are located in the project directory [`data_raw`](https://github.com/katekathrynkat/lizard-guts-naxos/data_raw/).

</div>

***

#### File locations

    data_raw
    ├── field                        # Field/lab measurements
        ├── original_spreadsheets        # Pre-cleaning data
            ├── [...].xlxs                  # Original Excel files
            └── data_errata.txt             # Log of corrections made to .csv files
        ├── diet_field.csv           # Stomach pump invert ID (all islands)
        ├── naxos_meta.csv           # Site metadata (Naxos)
        ├── naxos_veg.csv            # Vegetation quadrats (Naxos)
        ├── prey_meta.csv            # Pitfall/sticky metadata (all islands)
        ├── prey_pitfall.csv         # Pitfall trap invert ID (all islands)
        └── prey_sticky.csv          # Sticky trap invert ID (all islands)

Limitations:

- Prey/diet IDed to order
- Prey measured as relative abundance

```{r packages and data, class.source = 'fold-show'}

##### PACKAGES #####

library(tidyverse) # data manipulation & visualization

##### DATA #####

diet_raw <- read_csv("data_raw/field/diet_field.csv")
meta_raw <- read_csv("data_raw/field/naxos_meta.csv")
veg_raw <- read_csv("data_raw/field/naxos_veg.csv")
prey_meta_raw <- read_csv("data_raw/field/prey_meta.csv")
prey_p_raw <- read_csv("data_raw/field/prey_pitfall.csv")
prey_s_raw <- read_csv("data_raw/field/prey_sticky.csv")

```

## Field/lab data

### Site metadata

**Variables**

- `variable_name`: definition

```{r}
meta_raw
```

### Stomach pump invert ID

**Variables**

- `variable_name`: definition

```{r}
diet_raw
```

### Vegetation quadrats

**Variables**

- `variable_name`: definition

```{r}
veg_raw
```

### Pitfall/sticky metadata

**Variables**

- `variable_name`: definition

```{r}
prey_meta_raw
```

### Pitfall trap invert ID (all islands)

**Variables**

- `variable_name`: definition

```{r}
prey_p_raw
```

### Sticky trap invert ID (all islands)

**Variables**

- `variable_name`: definition

```{r}
prey_s_raw
```

<br>
<br>
<br>

<details><summary>**Session info**</summary>
```{r code}
devtools::session_info()
```
</details>

