---
title: "Data exploration"
subtitle: "From [Lizard guts: Naxos](https://github.com/katekathrynkat/lizard-guts-naxos)"
output:
  katereR::t_analysis
---

**Kate Culhane**, Dept. of Ecology, Evolution, and Marine Biology, UC Santa Barbara  
[kathrynculhane@ucsb.edu](kathrynculhane@ucsb.edu)
[<i class="fa fa-github"></i>](https://github.com/katekathrynkat/)
[<i class="fa fa-twitter"></i>](https://twitter.com/katekathrynkat/)

***

```{r setup, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(echo = TRUE, results = 'hold',
                      warning = FALSE, message = FALSE,
                      fig.width=5, fig.height=3, rows.print=5)
```

<div class = "summary">

**Summary**

Exploring the data

</div>

***

#### Dependencies

```{r packages and data, class.source = 'fold-show'}

##### PACKAGES #####

library(tidyverse) # data manipulation & visualization

##### DATA #####

# Metadata
meta <- read_csv("data_raw/field/naxos_meta.csv")

# Stomach pumping
diet <- read_csv("output/02_diet.csv")

# Available prey community
prey_meta <- read_csv("output/02_prey_meta.csv")
prey_p <- read_csv("output/02_prey_p.csv")
prey_s <- read_csv("output/02_prey_s.csv")

```

```{r}

# Data wrangling

prey_p_long <- prey_p %>% 
  pivot_longer(6:31, names_to = "order", values_to = "count")

prey_s_long <- prey_s %>% 
  pivot_longer(6:24, names_to = "order", values_to = "count")

prey <- full_join(prey_p_long, prey_s_long) %>% 
  full_join(meta[2:4])

```

## Prey community

```{r}

# Summary table

prey %>% 
  group_by(order) %>% 
  tally(count) %>% 
  arrange(-n)


prey %>% 
  group_by(order) %>% 
  tally(count) %>% 
  arrange(-n)

```


## Diet


## Notes & resources

Potential analyses:

- Diversity/abundance (prey and diet)
- Prey selectivity index
- NMDS
- Herbivory/myrmecophagy
- Selection for prey type
- Prey composition vs. habitat type / vegetation
- Biomass (using literature data)

To do: 

- Separate ants using comments

Data exploration (compare by land use category)

- Summarize prey/diet by frequency and incidence of each OTU
- Summarize abundance and diversity of prey/diet
- Summarize incidence of herbivory/myrmecophagy
- Flag rare OTUs
- NMDS of prey vs. diet

<br>
<br>
<br>

<details><summary>**Session info**</summary>
```{r code}
devtools::session_info()
```
</details>

